Predicting Severity of _C. difficile_ Infections from the Taxonomic Composition of the Gut Microbiome

### 2023-06-07 paper/paper.qmd
Link to cdiff relabun supp fig
Just show OTUs with any mean diff > 0.01

@@

::: {#fig-features fig-cap="**Feature importance.**
**A)** Feature importance via permutation test.
For each OTU, the order of samples was randomized in the test set 100 times and
the performance was re-calculated to estimate the permutation performance.
An OTU was considered important if the performance decreased when the OTU was
permuted in at least 75% of the models.
Mean difference in AUROC and the 75% confidence interval (CI) is reported for
each OTU, with starred OTUs being significant for the 75% CI.
OTUs with a greater difference in AUROC (actual performance minus permutation
performance) are more important.
Left: models were trained on the full datasets, with different numbers of
samples available for each severity definition.
Right: models were trained on the intersection of samples with all labels
available for each definition. Note that Attributable and Pragmatic severity are
exactly the same for the intersection dataset.
*Pseudomonas* (OTU 120) is not shown for IDSA severity in the full datasets nor
in the intersection dataset because it was removed during pre-processing due to
having near-zero variance.
**B)** Log~10~-transformed mean relative abundances of the most important OTUs
on the full datasets, grouped by severity (shape). 
The vertical dashed line is the limit of detection.
"}
![](figures/feature-importance.png)
:::

>>>>

knitr::opts_chunk$set(dpi = 600)

@@

::: {#fig-features fig-cap="**Most important OTUs for model performance.**
**A)** Feature importance via permutation test.
For each OTU, the order of samples was randomized in the test set 100 times and
the AUROC was re-calculated to estimate the permutation performance.
OTUs with a greater difference in AUROC (actual performance minus permutation
performance) are more important.
Mean difference in AUROC and the 75% confidence interval (CI) is reported for
each OTU that had a mean difference $\geq$ 0.01 for at least one severity
definition, with starred OTUs being significant for the 75% CI.
Notably, the OTU most likely corresponding to _C. difficile_ was not important
(see Supplementary @fig-cdiff).
Left: models were trained on the full datasets, with different numbers of
samples available for each severity definition.
Right: models were trained on the intersection of samples with all labels
available for each definition. Note that Attributable and Pragmatic severity are
exactly the same for the intersection dataset.
*Pseudomonas* (OTU 120) is not shown for IDSA severity in the full datasets nor
in the intersection dataset because it was removed during pre-processing due to
having near-zero variance.
**B)** Log~10~-transformed mean relative abundances of the most important OTUs
on the full datasets, grouped by severity (shape). 
The vertical dashed line is the limit of detection.
"}
![](figures/feature-importance.png)
:::
